---
title: "Global Natural Disaster Patterns from NASA EONET"
output:
  html_document:
    toc: false
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("process_eonet_data.R")
```

## Project Overview

This project analyzes global natural disaster observations from the NASA EONET API, with a focus on category imbalance, temporal trends, seasonality, spatial concentration, and predictive benchmarking.

## Main Research Question

How do global natural disaster observations in NASA EONET vary across category, time, seasonality, and space, and to what extent can these patterns support simple out-of-sample prediction?

## Key Findings

- Wildfires dominate the dataset and account for the large majority of observations.
- Recorded observations rise sharply in the early 2020s.
- Seasonal patterns differ across hazard categories.
- More flexible models provide limited overall gains over simple baselines.

## Featured Interactive Visualizations

### Annual Trend

```{r annual-trend, echo=FALSE}
p_annual
```

### Category-Specific Trends

```{r category-trend, echo=FALSE}
p_category_trend
```

### Spatial Distribution

```{r spatial-distribution, echo=FALSE}
p_map
```

## Explore More

- [Interactive Visualizations](Interactive_Vis.html)
- [Download Report (PDF)](final.pdf)
- [GitHub Repository](https://github.com/Rolland-He/JSC370-Project)
