---
title: "Global Natural Disaster Patterns from NASA EONET"
output:
  html_document:
    toc: false
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("process_eonet_data.R")
```

```{r summary-metrics, echo=FALSE}
top_cat <- category_counts$category[1]
top_cat_n <- category_counts$n[1]
top_cat_share <- category_counts$share[1]

peak_year <- annual_counts$year[which.max(annual_counts$n)]
peak_year_n <- max(annual_counts$n)

best_full_rmse <- model_eval %>% arrange(RMSE) %>% slice(1)
best_wf_rmse <- wf_model_eval %>% arrange(RMSE) %>% slice(1)
```

## Project Overview

This project analyzes global natural disaster observations from the NASA EONET API, with a focus on category imbalance, temporal trends, seasonality, spatial concentration, and predictive benchmarking.

## Main Research Question

How do global natural disaster observations in NASA EONET vary across category, time, seasonality, and space, and to what extent can these patterns support simple out-of-sample prediction?

## Key Findings

- `r top_cat` dominates the dataset with `r format(top_cat_n, big.mark = ",")` records (`r scales::percent(top_cat_share, accuracy = 0.1)`).
- Recorded observations peak in `r peak_year` at `r format(peak_year_n, big.mark = ",")`, with a sharp post-2023 rise.
- Seasonal structure is strong and category-dependent rather than uniform across hazards.
- In full-scope benchmarking, the best RMSE is from `r best_full_rmse$model` (`r round(best_full_rmse$RMSE, 3)`), while in wildfire-only benchmarking the best RMSE shifts to `r best_wf_rmse$model` (`r round(best_wf_rmse$RMSE, 3)`).

## Report Summary

The written report is structured as Introduction, Methods, Results, and Conclusions, and answers explicit research questions with quantitative evidence.

### Research question addressed in the report

- Main question: How do NASA EONET disaster observations vary by category, time, seasonality, and space, and how does predictive performance change between two targets (full multi-category counts vs wildfire-only counts)?
- Supporting questions include: category dominance, temporal trend shape, category-specific seasonality, and out-of-sample benchmark performance.

### Quantitative findings from the report

- Category imbalance is extreme: `r top_cat` accounts for `r scales::percent(top_cat_share, accuracy = 0.1)` of all observations.
- Temporal counts peak in `r peak_year` at `r format(peak_year_n, big.mark = ",")`; the surge is primarily wildfire-driven.
- Scope comparison changes model ranking: full scope best RMSE = `r round(best_full_rmse$RMSE, 3)` (`r best_full_rmse$model`), wildfire-only best RMSE = `r round(best_wf_rmse$RMSE, 3)` (`r best_wf_rmse$model`).
- Even with partial RMSE gains, out-of-sample R2 remains weak (negative in the report benchmarks), so predictive results are interpreted cautiously.

## Featured Interactive Visualizations

### Annual Trend

```{r annual-trend, echo=FALSE}
p_annual
```

### Category-Specific Trends

```{r category-trend, echo=FALSE}
p_category_trend
```

### Spatial Distribution

```{r spatial-distribution, echo=FALSE}
p_map_all
```

## Explore More

Explore category composition, seasonality panels, map filters, and predictive benchmarking on the full interactive page.

<a class="action-button" href="Interactive_Vis.html">Explore Full Interactive Visualizations</a>

- [Download Report (PDF)](final.pdf)
- [GitHub Repository](https://github.com/Rolland-He/JSC370-Project)
