---
title: "Interactive Visualizations"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
source("process_eonet_data.R")
```

## Category Dominance

```{r category-dominance, echo=FALSE}
p_category
```

This chart shows the category distribution of EONET observations.

## Temporal Trend

```{r temporal-trend, echo=FALSE}
p_annual
```

This line chart shows annual total observations over time.

## Category-Specific Trajectories

```{r category-trajectories, echo=FALSE}
p_category_trend
```

This interactive plot compares category-level annual trajectories.

## Seasonality

```{r seasonality, echo=FALSE}
p_heatmap
```

This heatmap shows month-by-category seasonal concentration.

## Spatial Distribution

```{r spatial-map, echo=FALSE}
p_map
```

The interactive map allows zooming and closer regional inspection.

## Report

For full methodology, statistical testing, predictive benchmarking, and limitations, see the full report.
