---
title: "Interactive Visualizations"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
source("process_eonet_data.R")
```

## Category Dominance

```{r category-dominance, echo=FALSE}
p_category
```

This ranked bar chart highlights the extreme category imbalance in the EONET record. Wildfires are visually emphasized and annotated because they dominate the observed event volume.

## Temporal Trends

```{r temporal-trend, echo=FALSE}
p_annual
```

The annual view combines raw counts with a smooth trend line to separate short-term fluctuations from broad movement in observation volume.

```{r category-trajectories, echo=FALSE}
p_category_trend
```

Use legend click and the chart controls to isolate Wildfires or non-wildfire categories and compare how the recent surge differs by hazard type.

## Seasonality {.tabset}

### Absolute Counts

```{r seasonality-abs, echo=FALSE}
p_heatmap_abs
```

This panel shows where absolute volume is concentrated by month and category.

### Within-Category Normalized

```{r seasonality-norm, echo=FALSE}
p_heatmap_norm
```

This normalized view rescales each category to reveal seasonal shape even for smaller categories.

### Monthly Curves by Category

```{r monthly-pattern-by-category, echo=FALSE}
p_monthly_pattern_by_category
```

Line-based monthly profiles complement the heatmap and make category-specific seasonality easier to trace.

## Spatial Distribution {.tabset}

### All Events

```{r map-all, echo=FALSE}
p_map_all
```

Clustered markers improve performance and make high-density regions easier to inspect.

### Wildfires Only

```{r map-wildfires, echo=FALSE}
p_map_wildfires
```

### Sea and Lake Ice Only

```{r map-ice, echo=FALSE}
p_map_ice
```

### Volcanoes Only

```{r map-volcanoes, echo=FALSE}
p_map_volcanoes
```

### Severe Storms Only

```{r map-storms, echo=FALSE}
p_map_storms
```

## Predictive Benchmarking

```{r model-compare, echo=FALSE}
p_model_compare
```

Use the metric toggle to switch among MAE, RMSE, and R2 for the four benchmark models.

`r xgboost_status_note`

```{r category-model-compare, echo=FALSE}
p_category_model_compare
```

Category-level RMSE comparison shows that gains are uneven and depend strongly on hazard type.

## Report

For full methodology, statistical testing, predictive benchmarking, and limitations, see the full report.
